---
title: "GrabFood Case"
author: "Patrick nguyen"
date: "2025-04-25"
output:
  html_document: default
  word_document: default
---
#IMPORTING LIBRARIES
```{r}
# 📦 Load all required packages
library(dplyr)
library(tidyr)
library(lubridate)
library(janitor)

library(ggplot2)
library(scales)
library(hrbrthemes)
library(ggrepel)
library(ggthemes)
library(patchwork)

library(readxl)
library(formattable)
library(gt)
library(DT)

library(broom)
library(car)
library(forecast)
library(tseries)

library(lpSolve)
library(treemapify)

```
### DEMAND FORECAST CASE
## PAT1:DATA CLEANING & FEATURE ENGINEERING
# 1.1 Load raw data
```{r}
df_raw <- read_excel("Case Study_ IMS Growth Specialist_Assessment.xlsx",sheet = "Assessment 1 - Forecasting",skip = 18)
```
# 1.2 Clean raw data
```{r}
df <- df_raw %>% 
  clean_names() %>% 
  rename(category = account)
# ⚙️ Create AOV inside df before summarizing
df <- df %>%
  mutate(
    AOV = if_else(booking > 0, gmv / booking, NA_real_)  # ✅ Create AOV (Average Order Value)
  )

```
# 1.3 Formt Numeric Column
```{r}
# Remove commas and convert fields to numeric
numeric_cols <- c("active_mex", "booking", "completed", 
                  "gmv", "grossbilling", "mex_spend", "grab_spend")

df <- df %>%
  mutate(across(all_of(numeric_cols), ~ as.numeric(gsub(",", "", as.character(.)))))

```
# 1.4 Create New Features
```{r}
# Feature engineering: new KPIs
df <- df %>%
  mutate(
    week = as.Date(week),
    month = month(week, label = TRUE),
    week_in_year = isoweek(week),
    week_in_month = ceiling(day(week) / 7),
    
    # Business Metrics
    total_spend = grab_spend + mex_spend,
    ROI = if_else(total_spend > 0, gmv / total_spend, NA_real_),
    CP_BO = if_else(booking > 0, total_spend / booking, NA_real_),
    CP_CO = if_else(completed > 0, total_spend / completed, NA_real_),
    Fulfillment_Rate = if_else(booking > 0, completed / booking, NA_real_),
    avg_booking_per_mex = if_else(active_mex > 0, booking / active_mex, NA_real_)
  )
```
##PART 2: WEEKLY DEMAND TREND ANALYSIS

# 2.1 Weekly Aggregation
```{r}
# 📊 Summarize weekly performance
weekly_trend <- df %>%
  group_by(week, week_in_year) %>%
  summarise(
    Booking = sum(booking, na.rm = TRUE),
    Spend = sum(total_spend, na.rm = TRUE),
    GMV = sum(gmv, na.rm = TRUE),
    ROI = sum(gmv, na.rm = TRUE) / sum(total_spend, na.rm = TRUE),
    avg_booking_per_mex = mean(avg_booking_per_mex, na.rm = TRUE),
    Fulfillment_Rate = mean(Fulfillment_Rate, na.rm = TRUE),
    CP_BO = mean(CP_BO, na.rm = TRUE),
    CP_CO = mean(CP_CO, na.rm = TRUE),
    AOV = mean(AOV, na.rm = TRUE),        # ✅ Now safe to summarize AOV
    .groups = "drop"
  ) %>%
  mutate(
    Spend = round(Spend, 0),
    GMV = round(GMV, 0),
    Booking = round(Booking, 0),
    ROI = round(ROI, 2),
    avg_booking_per_mex = round(avg_booking_per_mex, 2),
    Fulfillment_Rate = round(Fulfillment_Rate, 2),
    CP_BO = round(CP_BO, 2),
    CP_CO = round(CP_CO, 2),
    AOV = round(AOV, 2)
  )
```
#2.2 Plot Combo Chart to spot trend
```{r}
# 👇 Convert week_in_year to numeric for proper timeline spacing
weekly_trend <- weekly_trend %>%
  mutate(week_numeric = as.numeric(week_in_year))

# Recalculate scaling factor (still needed)
scaling_factor <- max(weekly_trend$Spend, na.rm = TRUE) / max(weekly_trend$ROI, na.rm = TRUE)

# ✅ Updated ggplot with correct week axis
ggplot(weekly_trend, aes(x = week_numeric)) +

  # 1️⃣ Area Chart for Booking Orders
  geom_area(aes(y = Booking, fill = "Booking Order"), alpha = 0.5) +

  # 2️⃣ Bar Chart for Total Spend
  geom_col(aes(y = Spend, fill = "Total Spend"), alpha = 0.8) +

  # 3️⃣ Line Chart for ROI (scaled)
  geom_line(aes(y = ROI * scaling_factor, color = "ROI"), size = 1.5) +

  # 4️⃣ ROI Labels
  geom_text(aes(y = ROI * scaling_factor + 10000, label = round(ROI, 2)), 
            color = "darkgrey", size = 4, fontface = "bold") +

  # 5️⃣ Trend Lines
  geom_smooth(aes(y = Booking, color = "Booking Order Trend"), method = "lm", se = FALSE, linetype = "dotted") +
  geom_smooth(aes(y = Spend, color = "Total Spend Trend"), method = "lm", se = FALSE, linetype = "dotted") +

  # 🏷 Labels
  labs(
    title = "Demand Trend by Week",
    x = "Week in Year",
    y = NULL,
    caption = "ROI is shown on the secondary axis\nDotted lines represent linear trends"
  ) +

  # 🎨 Colors
  scale_fill_manual(values = c("Booking Order" = "yellow", "Total Spend" = "black")) +
  scale_color_manual(values = c(
    "ROI" = "orange",
    "Booking Order Trend" = "red",
    "Total Spend Trend" = "green"
  )) +

  # 🪄 Fix Y-axis + Dual ROI axis
  scale_y_continuous(
    labels = scales::comma,
    sec.axis = sec_axis(~ . / scaling_factor, name = NULL, labels = comma)
  ) +

  # ✅ Fix X-axis: show each week cleanly
  scale_x_continuous(
    breaks = weekly_trend$week_numeric,
    labels = weekly_trend$week_in_year
  ) +

  # ✨ Clean theme
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5),  # 👉 Weeks stand upright
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  )
```
## PART3: L3W Deepdive by Segment Analysis
#3.1 Filter Data to Weeks 47-49
```{r}
# 📚 Load libraries
library(dplyr)
library(scales)
library(formattable)

# 📊 Step 1: Create main pivot without subtotal
pivot_main <- df %>%
  filter(week_in_year %in% c(47, 48, 49)) %>%
  select(category, week_in_year, active_mex, total_spend, booking) %>%
  arrange(category, week_in_year) %>%
  group_by(category) %>%
  mutate(
    pct_change_active_mex = (active_mex / lag(active_mex) - 1) * 100,
    pct_change_spent = (total_spend / lag(total_spend) - 1) * 100,
    pct_change_booking = (booking / lag(booking) - 1) * 100,
    avg_booking_per_mex = booking / active_mex,
    pct_change_avg_demand_mex = (avg_booking_per_mex / lag(avg_booking_per_mex) - 1) * 100
  ) %>%
  mutate(across(starts_with("pct_change"), ~ replace_na(.x, 0))) %>%
  rename(Time = week_in_year) %>%
  mutate(Time = factor(Time, levels = c(47, 48, 49))) %>%
  ungroup()

# 📊 Step 2: Split by category and add subtotal after each category
pivot_list <- pivot_main %>%
  group_split(category)

pivot_final <- lapply(pivot_list, function(group_df) {
  subtotal <- group_df %>%
    summarise(
      category = unique(category),
      Time = "Subtotal",
      active_mex = sum(active_mex, na.rm = TRUE),
      total_spend = sum(total_spend, na.rm = TRUE),
      booking = sum(booking, na.rm = TRUE),
      avg_booking_per_mex = sum(booking, na.rm = TRUE) / sum(active_mex, na.rm = TRUE),
      pct_change_active_mex = NA,
      pct_change_spent = NA,
      pct_change_booking = NA,
      pct_change_avg_demand_mex = NA,
      .groups = "drop"
    )
  bind_rows(group_df, subtotal)
}) %>%
  bind_rows() %>%
  select(
    category, Time,
    active_mex, total_spend, booking, avg_booking_per_mex,
    pct_change_active_mex, pct_change_spent, pct_change_booking, pct_change_avg_demand_mex
  )

# 📋 Step 3: Format number and percentage
pivot_table <- pivot_final %>%
  mutate(
    active_mex = comma(round(active_mex, 0)),
    total_spend = comma(round(total_spend, 0)),
    booking = comma(round(booking, 0)),
    avg_booking_per_mex = round(avg_booking_per_mex, 2),
    pct_change_active_mex = sprintf("%.1f%%", pct_change_active_mex),
    pct_change_spent = sprintf("%.1f%%", pct_change_spent),
    pct_change_booking = sprintf("%.1f%%", pct_change_booking),
    pct_change_avg_demand_mex = sprintf("%.1f%%", pct_change_avg_demand_mex)
  )

# 🎨 Step 4: Define color scales
color_scale_active_mex <- function(x) {
  color_tile("yellow", "green")(as.numeric(x))
}
color_scale_spent <- function(x) {
  color_tile("green", "yellow")(as.numeric(x))
}
color_scale_booking <- function(x) {
  color_tile("yellow", "green")(as.numeric(x))
}
color_scale_avg_demand_mex <- function(x) {
  color_tile("yellow", "green")(as.numeric(x))
}

# 🎯 Step 5: Apply color formatting
pivot_table_formatted <- pivot_table %>%
  group_by(category) %>%
  mutate(
    pct_change_active_mex = color_scale_active_mex(gsub("%", "", pct_change_active_mex)),
    pct_change_spent = color_scale_spent(gsub("%", "", pct_change_spent)),
    pct_change_booking = color_scale_booking(gsub("%", "", pct_change_booking)),
    pct_change_avg_demand_mex = color_scale_avg_demand_mex(gsub("%", "", pct_change_avg_demand_mex))
  ) %>%
  ungroup()

# 📋 Step 6: Display final pivot heatmap
formattable(pivot_table_formatted)

```
#3.2 Export the heatmap into PNG
```{r}
install.packages("webshot2")

library(formattable)
library(htmlwidgets)
library(webshot2)  # Install with: install.packages("webshot2")

# Save the HTML widget to a temporary HTML file
saveWidget(as.htmlwidget(formattable(pivot_table_formatted)), "pivot_heatmap.html", selfcontained = TRUE)

# Take a screenshot of the HTML file and save as PNG
webshot("pivot_heatmap.html", 
        "pivot_heatmap_fixed.png", 
        zoom = 2, 
        vwidth = 1600, 
        vheight = 1800)  # Increase height significantly


```


## PART4: Multiple Regression Model
# 4.1 Prepare the Clean Dataset for Regression
```{r}
# 🧹 Step 1.1: Select necessary columns for modeling
regression_data <- df %>%
  select(category, week_in_year, booking, total_spend, active_mex, ROI, Fulfillment_Rate) %>%
  filter(!is.na(booking), !is.na(total_spend), !is.na(active_mex))  # Remove missing rows if any

# 🏷️ Step 1.2: Create Dummy Variables for Category
regression_data <- regression_data %>%
  mutate(
    A2_dummy = if_else(category == "A2", 1, 0),
    A3_dummy = if_else(category == "A3", 1, 0),
    A4_dummy = if_else(category == "A4", 1, 0)
    # (A1 is the baseline, no dummy needed)
  )

# 🧹 Step 1.3: Optional — Check glimpse of the cleaned dataset
glimpse(regression_data)

```
# 4.2 Correlation Matrix
```{r}
# 📊 Step 2.1: Select only numerical columns for correlation
correlation_data <- regression_data %>%
  select(booking, total_spend, active_mex, ROI, Fulfillment_Rate)

# 📈 Step 2.2: Compute Correlation Matrix
correlation_matrix <- cor(correlation_data, use = "complete.obs")

# 🖼️ Step 2.3: Visualize Correlation Matrix
library(corrplot)

# Basic Corrplot
corrplot(
  correlation_matrix,
  method = "color",       # Show color tiles
  type = "upper",         # Only upper half
  addCoef.col = "black",  # Add correlation coefficients
  number.cex = 0.7,       # Size of correlation numbers
  tl.col = "black",       # Text labels color
  tl.srt = 45             # Rotate labels
)

```
#cmt>> multicollinear of active_mex & spending >> take out

#4.3 Finetune Regression Model with catagorical variables
# Regression Formula: booking ~ total_spend + A2 + A3 + A4
```{r}
# 📚 Load necessary libraries
library(dplyr)
library(broom)

# ⚙️ Step 1: Prepare the clean dataset
# Assume `df` is your main dataset already cleaned previously

# Create dummy variables for A2, A3, A4 (A1 is baseline)
df_model <- df %>%
  filter(week_in_year %in% c(47, 48, 49)) %>%  # Focus on L3W (week 47-49)
  mutate(
    A2 = if_else(category == "A2", 1, 0),
    A3 = if_else(category == "A3", 1, 0),
    A4 = if_else(category == "A4", 1, 0)
  ) %>%
  select(booking, total_spend, A2, A3, A4)

# ✅ Step 2: Build the Regression Model
model_booking <- lm(booking ~ total_spend + A2 + A3 + A4, data = df_model)

# 📋 Step 3: Output the Summary
summary(model_booking)

# 📋 Step 4: Optional — Pretty tidy summary for report
model_booking_tidy <- broom::tidy(model_booking)
print(model_booking_tidy)

# 📊 Step 5: Optional — Check R-squared, Adjusted R-squared
model_stats <- broom::glance(model_booking)
print(model_stats)

# 📍 Interpretation Hints:
# - Coefficients: Positive/negative impact
# - p-value < 0.05 = statistically significant impact
# - R-squared: Overall model fit

# 🚀 Now we can interpret: Which category (A2/A3/A4) matters most!

```

## PART5: Forecasting Model for 6 weeks (AIRMA & Econometric)
#5.1 Forecast Booking based on 2 model (econometrics is more reliable)
```{r}
# 📚 Load libraries
library(dplyr)
library(forecast)
library(tidyr)
library(broom)
library(ggplot2)

# ⚙️ Assume your data 'df_model' already has:
# Columns: booking, total_spend, A2, A3, A4

# ==========================================================
# 1️⃣ ECONOMETRIC FORECASTING (Regression-based Prediction)
# ==========================================================

# Fit regression model
model_reg <- lm(booking ~ total_spend + A2 + A3 + A4, data = df_model)

# View model summary
summary(model_reg)

# Forecast next 6 weeks:
# (You need to simulate future spend and A2, A3, A4 flags)

# Example: Create new input assumptions
future_input <- data.frame(
  total_spend = c(1500000, 1550000, 1600000, 1650000, 1700000, 1750000),   # Forecasted spend
  A2 = c(0, 0, 0, 0, 0, 0),   # Category flags
  A3 = c(1, 1, 1, 1, 1, 1),
  A4 = c(0, 0, 0, 0, 0, 0)
)

# Predict Booking for future
booking_forecast_reg <- predict(model_reg, newdata = future_input)

# Create final Econometric Forecast Table
econometric_forecast <- future_input %>%
  mutate(
    forecasted_booking = round(booking_forecast_reg)
  )

print(econometric_forecast)

# ==========================================================
# 2️⃣ TIME SERIES FORECASTING (ARIMA model)
# ==========================================================

# Create time series object from booking
booking_ts <- ts(df_model$booking, frequency = 1)

# Fit ARIMA model automatically
model_arima <- auto.arima(booking_ts)

# View model summary
summary(model_arima)

# Forecast 6 steps ahead
forecast_arima <- forecast(model_arima, h = 6)

# Convert to dataframe
arima_forecast <- data.frame(
  Week = seq_len(6),
  Forecasted_Booking = round(forecast_arima$mean),
  Lower_80 = round(forecast_arima$lower[,1]),
  Upper_80 = round(forecast_arima$upper[,1])
)

print(arima_forecast)

# ==========================================================
# 3️⃣ VISUALIZE COMPARISON
# ==========================================================

# Combine both results
combined_forecast <- data.frame(
  Week = seq(1,6),
  Booking_Econometric = econometric_forecast$forecasted_booking,
  Booking_ARIMA = arima_forecast$Forecasted_Booking
)

# Plot
ggplot(combined_forecast, aes(x = Week)) +
  geom_line(aes(y = Booking_Econometric, color = "Econometric"), size = 1.2) +
  geom_line(aes(y = Booking_ARIMA, color = "ARIMA"), size = 1.2, linetype = "dashed") +
  geom_point(aes(y = Booking_Econometric, color = "Econometric"), size = 2) +
  geom_point(aes(y = Booking_ARIMA, color = "ARIMA"), size = 2) +
  labs(
    title = "Booking Forecast: Econometric vs ARIMA",
    x = "Future Week",
    y = "Booking Orders",
    color = "Model"
  ) +
  theme_minimal(base_size = 14)

```
#5.2 Forcecast for GMV, spend, ROI (case base without budget reallocation)
```{r}
# 📚 Load libraries
library(dplyr)
library(ggplot2)

# 📊 Step 1: Assume Last Week Spend by Category
spend_last_week <- tibble(
  category = c("A1", "A2", "A3", "A4"),
  spend = c(370000, 320000, 330000, 380000)  # Example values
)

# 📋 Step 2: Reallocate Budget based on New Strategy
reallocation_plan <- tibble(
  category = c("A1", "A2", "A3", "A4"),
  adjustment = c(-0.10, -0.03, 0.15, -0.10)  # Cut or Boost
)

# Calculate new spend
spend_reallocated <- spend_last_week %>%
  left_join(reallocation_plan, by = "category") %>%
  mutate(new_spend = spend * (1 + adjustment))

# Normalize to make total spend same as original
total_original_spend <- sum(spend_last_week$spend)
total_new_spend <- sum(spend_reallocated$new_spend)
scaling_factor <- total_original_spend / total_new_spend

spend_reallocated <- spend_reallocated %>%
  mutate(final_spend = new_spend * scaling_factor)

# 🧮 Step 3: Assumptions for AOV, CPBO, Fulfillment Rate
future_plan <- tibble(
  week = 1:6,
  AOV = c(12.0, 12.2, 12.3, 12.4, 12.5, 12.6),  # Example slight growth
  CPBO = c(0.90, 0.92, 0.93, 0.94, 0.95, 0.96),
  fulfillment_rate = c(0.92, 0.92, 0.91, 0.91, 0.91, 0.90)
)

# 🏗 Step 4: Forecast Booking Orders
# Regression coefficients (from your earlier regression)
coef_spend <- -0.105
coef_A2 <- -136400
coef_A3 <- 461500
coef_A4 <- -454700

# Assume we boost A3 and reduce others as per reallocated budget
future_forecast <- future_plan %>%
  mutate(
    total_spend = total_original_spend,  # Assume total spend kept same
    # Booking from regression model
    booking = (coef_spend * total_spend) + 
              (coef_A2 * 0) +   # Assume A2 kept neutral
              (coef_A3 * 1) +   # Push A3
              (coef_A4 * 0) +   # Cut A4
              491500,           # Intercept
    booking = pmax(booking, 0)  # Avoid negative booking
  )

# Step 5: Forecast GMV, Completed Orders, ROI
future_forecast <- future_forecast %>%
  mutate(
    GMV = booking * AOV,
    completed_orders = booking * fulfillment_rate,
    ROI = GMV / total_spend
  )

# 📈 Step 6: Plot GMV and ROI Forecast
p <- ggplot(future_forecast, aes(x = week)) +
  geom_line(aes(y = GMV/1e6, color = "GMV ($M)"), size = 1.2) +
  geom_line(aes(y = ROI * 5, color = "ROI (scaled)"), linetype = "dashed", size = 1.2) +
  scale_y_continuous(
    name = "GMV ($M)",
    sec.axis = sec_axis(~./5, name = "ROI")
  ) +
  scale_color_manual(values = c("GMV ($M)" = "blue", "ROI (scaled)" = "red")) +
  labs(
    title = "Forecast under Reallocation toward A3",
    x = "Week",
    color = "Metric"
  ) +
  theme_minimal()

# ✨ Output Forecast Table and Plot
print(future_forecast)
print(p)

```
#5.3 Adjust Forcecast for GMV, spend, ROI (based on domain knwoledge of changes range in real life business)
```{r}
# 📦 Load necessary libraries
library(dplyr)
library(tibble)

# 📊 1. Baseline Fix Scenario
baseline_fix <- tibble(
  week = 50:55,
  booking = rep(806000, 6),
  AOV = c(12.0, 12.2, 12.3, 12.4, 12.5, 12.6),
  CPBO = c(0.90, 0.92, 0.93, 0.94, 0.95, 0.96),
  fulfillment_rate = c(0.92, 0.92, 0.91, 0.91, 0.91, 0.90),
  total_spend = rep(1400000, 6)
) %>%
  mutate(
    completed_orders = round(booking * fulfillment_rate),
    GMV = round(booking * AOV),
    ROI = round(GMV / total_spend, 2)
  )

# 📊 2. Optimized Reallocation Scenario
optimized <- tibble(
  week = 50:55,
  booking = rep(832000, 6),  # Adjusted upward from regression result
  AOV = c(12.0, 12.2, 12.3, 12.4, 12.5, 12.6),
  CPBO = c(0.90, 0.92, 0.93, 0.94, 0.95, 0.96),
  fulfillment_rate = c(0.92, 0.92, 0.91, 0.91, 0.91, 0.90),
  total_spend = rep(1400000, 6)
) %>%
  mutate(
    completed_orders = round(booking * fulfillment_rate),
    GMV = round(booking * AOV),
    ROI = round(GMV / total_spend, 2)
  )

# ✅ View Results
cat("📘 Baseline Fix:\n")
print(baseline_fix)

cat("\n📗 Optimized Reallocation:\n")
print(optimized)

```
#5.4 Final forecast model & result
```{r}
# 📚 Load necessary libraries
library(dplyr)
library(formattable)

# 📊 Historical spend and assumptions (Weeks 46–49)
historical_spend <- c(1537401, 1594337, 1709213, 1759203)

# 🔧 Build baseline scenario using historical spend
baseline_forecast <- data.frame(
  week = 50:55,
  total_spend = c(historical_spend, rep(mean(historical_spend), 2)),
  AOV = c(12.0, 12.2, 12.3, 12.4, 12.5, 12.6),
  CPBO = c(0.90, 0.92, 0.93, 0.94, 0.95, 0.96),
  fulfillment_rate = c(0.92, 0.92, 0.91, 0.91, 0.91, 0.90)
)

baseline_forecast <- baseline_forecast %>%
  mutate(
    booking = round(total_spend / CPBO),
    completed_orders = round(booking * fulfillment_rate),
    GMV = round(completed_orders * AOV),
    ROI = round(GMV / total_spend, 2)
  )

# 🌱 Optimized scenario: Assume same spend but better allocation (+3% uplift in booking)
optimized_forecast <- baseline_forecast %>%
  mutate(
    booking = round(booking * 1.03),
    completed_orders = round(booking * fulfillment_rate),
    GMV = round(completed_orders * AOV),
    ROI = round(GMV / total_spend, 2)
  )
# Add scenario label to each dataset
baseline_forecast <- baseline_forecast %>%
  mutate(Scenario = "Baseline Fix")

optimized_forecast <- optimized_forecast %>%
  mutate(Scenario = "Optimized Reallocation")

# Print each table clearly
cat("\n📘 Baseline Forecast Table:\n")
print(baseline_forecast)

cat("\n📗 Optimized Reallocation Forecast Table:\n")
print(optimized_forecast)


```

### PROMOTION CASE
## PART6: Promotion Scenario Optimization Model
## PART7: Decision Making Model
















































































